Coherence of Molecular Mechanisms In Major Human Disease and Traits

### 2020-02-10 manuscript/submission.BMC_Bioinformatics/manuscript_BMC_Bioinformatics_R1.Rmd
Start revision by adding modularity results/discussion/methods description

@@

<!--
## Single networks, or networks of modules?

Our study considers a network of disease-associated genes in its entirety. However, networks may be comprised of interacting modules, i.e., relatively isolated subnetworks [@Hartwell:1999aa]. Connectivity of such modules would be different from the overall network connectivity [@Ghiassian:2015aa]. Although numerous methods exist for module detection in a network [@Chen:2008aa; @Dobrin:2009aa; @Jia:2011aa; @Agrawal:2018aa], comparing modules between diseases is plagued by normalization issues described in the Results section Intuitively, large networks will have a higher chance to have subnetworks (modules), and vice versa. It remains to be determined whether disease-associated genes form modular networks and whether the modular structure differs between (categories of) diseases.

## Tissue specificity

>>>>

<!--New start-->
## Coherence captures distinct network properties than modularity

Our study considers a network of phenotype-associated genes in its entirety. However, networks may be comprised of interacting modules, i.e., relatively isolated subnetworks [@Hartwell:1999aa]. Such subnetworks, while highly coherent by themselves, would have relatively low coherence overall due to the limited number of interactions among them (canonical definition of subnetwork). Consequently, a network with low coherence may consist of several modules and have high modularity. In turn, a network with high coherence may be represented by a single module and, consequently, have low modularity. Our assessment of network modularity using STRING and STRING filtered PPIs revealed negative association of modularity with network size, that is, the larger the network is, the less modular it becomes (Additional File 3, 4). We also did not observe a measurable association between modularity and coherence. Notably, network modularity using BioGrid PPIs showed a moderate positive association with network size and coherence, implying that, as networks get larger, they become more modular and yet highly coherent. Investigation of individual phenotype-specific networks built using BioGrid PPIs revealed they are expectedly smaller than networks build using STRING PPIs, making modularity estimates imprecise. In summary, our results did not reveal a consistent association between coherence and modularity, suggesting these metrics capture distinct network properties.

Although our results did not show an association between coherence and modularity, our study is limited by using a single most common modularity measure [@Clauset_2004]. Although numerous methods for module detection have been described [@Chen:2008aa; @Dobrin:2009aa; @Jia:2011aa; @Agrawal:2018aa], comparing modules between diseases is plagued by normalization issues described in the Results section. Our future direction includes the development of a strategy for normalization of modularity by network size and comparing it with normalized coherence.

## Modularity estimates

Network modularity was estimated using the standard method developed by Clauset et al. [@Clauset_2004], implemented in the igraph::modularity function.
<!--New end-->

@@

<!--
## Tissue specificity
